---
title: "Stochastic Streamflow Generation: A knn approach under a changing climate"
author: "Ze Jiang"
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2:
  #bookdown::pdf_document2:
    toc: true
    highlight: tango
    number_section: false
geometry: "left=3cm,right=3cm,top=2cm,bottom=2cm"
knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = file.path(dirname(inputFile), paste0('ALL_GCM_r100_v1'))) }) #,format(Sys.time(),"%Y-%m-%d %H %M")))) })
editor_options: 
  chunk_output_type: console
---

# Setup

```{r setup, include=TRUE}
# this could remove the params
# rm(list=ls()); graphics.off() # clear environment and graphics
library(knitr)
#devtools::install_github(repo="haozhu233/kableExtra", ref="a6af5c0")
library(kableExtra)
library(rmarkdown)

# Global parameters
op <- par() # save original par

nensemble <- 100
flag.sel <- switch(1, "ALL","NPRED","WASP"); flag.sel
flag.ver <- switch(2, "","_v1","_v2")

# Demo station
station_id <- "Q5"

#flag.save <- F
font_family <- "serif"
size_base <- 16
size_text <- 20

# Places the float at precisely the location in the pandoc code. 
# Requires the float package. "H" is somewhat equivalent to "h!".
knitr::opts_chunk$set(echo = TRUE, include=TRUE, collapse = TRUE, comment = "#>", 
                      message = FALSE, warning = FALSE,
                      
                      #fig.path = paste0("figure/","-",flag.sel,flag.ver,"-"), 
                      dev = "jpeg", dpi=500, out.width = '85%',
          					  fig.height = 9, fig.width = 9,
          					  fig.align='center', fig.pos="h!")

```

# Required packages

```{r packages, echo=TRUE, warning=FALSE}
library(Qgen)

library(lubridate)
library(moments)
library(ggplot2) 
library(zoo)
library(dplyr)
library(tidyr)
library(scales)
library(data.table)

require(ggpubr) # ggplot2
require(egg) # facet
require(lemon) # facet

```

# Load data

Observed streamflow (Q) and its conditional variables, Precipitation (P), Potential ET(PET), Temperature (T) as well as the water blance (P-PET). 

```{r dat, message=FALSE}
# obsvered Q
data("Harz_obs_Q")

Qday_obs <- Harz_obs_Q[[station_id]] %>% rename("year"="iyear","month"="imon","day"="iday","Qd"="Qval")
Qmon_obs <- aggregate(Qd~year+month, Qday_obs, FUN = mean) %>% rename("Qm"="Qd") %>% arrange(year,month)
Qyr_obs <- stats::aggregate(Qd~year, Qday_obs, FUN = mean) %>% rename("Qa"="Qd") %>% arrange(year)

head(Qmon_obs[,1:2])
tail(Qmon_obs[,1:2])
date_Qobs <- as.Date(paste(Qmon_obs[,1],Qmon_obs[,2],"1", sep="-"))
range(date_Qobs)

# simulated monthly & daily
path.out <- paste0(station_id,"/") 
load(paste0(path.out, "Harz_Qmon_",flag.sel,"_r",nensemble,"_hist",flag.ver,".Rdat")) # Qsim_mon
load(paste0(path.out, "Harz_Qday_",flag.sel,"_r",nensemble,"_hist",flag.ver,".Rdat")) # Qsim_day
summary(Qsim_mon[[1]])

# merage
Qmon_obs1 <- Qmon_obs %>% rename("obs"="Qm")
Qmon_sim <- rbindlist(Qsim_mon, idcol="group") %>% rename("sim"="Qm") %>% 
    merge(Qmon_obs1, by=c("year","month"))  %>% 
    mutate(season=cut(month, breaks=c(0, 3, 6, 9, 12), right = T, labels = c(1,2,3,4)))

# dates of hist ----
val.st <- 1971; val.end <- 2000
date_year <- val.st:val.end 
date_year - unique(Qsim_mon[[1]]$year) #%>% range() # cross check

date_mon <- seq(as.Date(paste(val.st,"1","1",sep="-")),
                  as.Date(paste(val.end,"12","1",sep="-")),
                  by="month")

date_day <- seq(as.Date(paste(val.st,"1","1",sep="-")),
                as.Date(paste(val.end,"12","31",sep="-")),
                by="day")


ind_yrs <- which(unique(Qmon_obs[,1]) %in% date_year)
ind_mon <- which(date_Qobs %in% date_mon)
ind_day <- which(paste0(Qday_obs[,1], Qday_obs[,2], Qday_obs[,3]) %in%
                   paste0(year(date_day), month(date_day), day(date_day)))
```

# Monthly

## Caculate statistics

```{r mon}
#------------------------------------------------------------------------#
# OBS input
Qmon=Qmon_obs$Qm[ind_mon]
mon=Qmon_obs$mon[ind_mon]
year=Qmon_obs$year[ind_mon]

# SIM input
Qmon1 <- sapply(Qsim_mon, function(ls) ls$Qm)
#summary(Qmon1)
mon1=mon
year1=year

plot(date_mon, Qmon, type='l', col="white")
for(i_ens in 1:3){
  lines(date_mon, Qmon1[,i_ens],  col="grey")
}
lines(date_mon, Qmon, type='l', col="red")


nyr_obs <- val.end-val.st+1
nmo_obs <- nyr_obs*12
nyr_sim <- nyr_obs
nmo_sim <- nmo_obs
nrea <- nensemble
#------------------------------------------------------------------------#
# OBS initialisations
Qm <- numeric(length=12)    # mean
Qs <- numeric(length=12)    # stddev
Qmin <- numeric(length=12)  # min
Qmax <- numeric(length=12)  # max
Qsk <- numeric(length=12)  # skewness
Qcor <- numeric(length=12)  # lag-1 autocorrelation
Qacf_su <- numeric(length=20)  # ACF summer 20 lags
Qacf_wi <- numeric(length=20)  # ACF winter 20 lags

# SIM initialisations
Qm1 <- array(NA, dim=c(nrea,12))    # mean (dim: 'nrea' realisations, 12 months)
Qs1 <- array(NA, dim=c(nrea,12))    # stddev
Qmin1 <- array(NA, dim=c(nrea,12))  # min
Qmax1 <- array(NA, dim=c(nrea,12))  # max
Qsk1 <- array(NA, dim=c(nrea,12))   # skewness
Qcor1 <- array(NA, dim=c(nrea,12))   # lag-1 autocorrelation
Qacf1_su <- array(NA, dim=c(nrea,20))   # ACF summer 20 lags
Qacf1_wi <- array(NA, dim=c(nrea,20))   # ACF summer 20 lags
amin <- numeric(length=20)  # min ACF 20 lags
amax <- numeric(length=20)  # max ACF 20 lags
VDmax1 <- numeric(length=nrea)  # max deficit volume (1)
DDmax1 <- numeric(length=nrea)  # max deficit duration (2)
VSmax1 <- numeric(length=nrea)  # max surplus volume (3)
DSmax1 <- numeric(length=nrea)  # max surplus duration (4)
SM1 <- numeric(length=nrea)     # storage capacity for full compensation (5)
LDP1 <- numeric(length=nrea)    # longest dry period for full compensation (6)
Diff1 <- array(NA, dim=c(nyr_sim*12,nrea))   # mass curve cumsum(Q-MQ); 'nrea' realis; 
sc <- array(NA, dim=c(nrea,6))    # storage characteristics (dim: 'nrea' realisations, 6 characteristics)

#------------------------------------------------------------------------#
# Basic OBS statistics by month
for (i in 1:12) {
  Qm[i] <- mean(Qmon[mon == i])
  Qs[i] <- sd(Qmon[mon == i])
  Qmin[i] <- min(Qmon[mon == i])
  Qmax[i] <- max(Qmon[mon == i])
  Qsk[i] <- skewness(Qmon[mon == i])  
}

# Basic SIM statistics by month for each realization
#summary(Qmon1)
for (i in 1:12) {
  for (j in 1:nrea) {
    Qm1[j,i] <- mean(Qmon1[,j][mon1 == i])
    Qs1[j,i] <- sd(Qmon1[,j][mon1 == i])
    Qmin1[j,i] <- min(Qmon1[,j][mon1 == i])
    Qmax1[j,i] <- max(Qmon1[,j][mon1 == i])
    Qsk1[j,i] <- skewness(Qmon1[,j][mon1 == i]) 
  }
}

# Lag-1 autocorrelation per month for OBS & SIM
Qcor <- ac1mon(matrix(Qmon, ncol=12, byrow=T))
for (j in 1:nrea) { Qcor1[j,1:12]=ac1mon(matrix(Qmon1[,j], ncol=12,byrow=T))}

# ACF OBS & SIM 20 lags for summer and winter
Qacf_su <- acf(Qmon[mon>4 & mon<11], plot=FALSE)$acf[2:16]    # OBS summer
Qacf_wi <- acf(Qmon[mon<5 | mon>10], plot=FALSE)$acf[2:16]    # OBS winter
for(j in 1:nrea) { Qacf1_su[j,1:15] <- acf(Qmon1[,j][mon1>4 & mon1<11], plot=FALSE)$acf[2:16] }    # SIM nreas summer
for(j in 1:nrea) { Qacf1_wi[j,1:15] <- acf(Qmon1[,j][mon1<5 | mon1>10], plot=FALSE)$acf[2:16] }    # SIM nreas winter
```

## Basic statistics - monthly

```{r fig-mon-stat}
#if(flag.save) jpeg("Qmon_valid.jpg", units="cm", width=16.5, height=22, res=dpi_fig)

# plot layout as matrix filled by cols
layout(matrix(c(1:10), nrow=5, ncol=2, byrow=TRUE))    
par(ps="12",              # size
    las=1,                # axis label orientation
    mar=c(2,5,1,1),       # no. of plot margin lines
    oma=c(2,2,2,1),       # no. of outer margin lines
    mgp=c(2.5,0.5,0),         # margin line for the axis title, labels and line.
    pty="m")              # maximum plot region used

# Box-Plot mean monthly Q
boxplot(Qm1, range=0, col = "lightblue", xlab='Month', ylab='Average [m3/s]')   # grouped by month
points(Qm,pch=4,col="red",lwd=2)         # add obs mean value as red cross

# Box-Plot StDev monthly Q
boxplot(Qs1, range=0, col = "lightblue", xlab='Month', ylab='Std. Dev. [m3/s]')   # grouped by month
points(Qs,pch=4,col="red",lwd=2)         # add obs mean value as red cross

# Box-Plot Min monthly Q
boxplot(Qmin1, range=0, col = "lightblue", xlab='Month', ylab='Min [m3/s]')   # grouped by month
points(Qmin,pch=4,col="red",lwd=2)         # add obs mean value as red cross

# Box-Plot Max monthly Q
boxplot(Qmax1, range=0, col = "lightblue", xlab='Month', ylab='Max [m3/s]')   # grouped by month
points(Qmax,pch=4,col="red",lwd=2)         # add obs mean value as red cross

# Box-Plot Schiefe monthly Q
boxplot(Qsk1, range=0, col = "lightblue", xlab='Month', ylab='Schiefe [m3/s]')   # grouped by month
points(Qsk,pch=4,col="red",lwd=2)         # add obs mean value as red cross

# Box-Plot Lag-1 autocorrelation monthly Q
boxplot(Qcor1, range=0, col = "lightblue", xlab='Month', ylab='Lag-1 Corr [-]')   # grouped by month
points(Qcor,pch=4,col="red",lwd=2)         # add obs mean value as red cross
abline(a=0, b=0, col="black",lty=2)

# Calculate storage statistics for OBS flows----------------------------------->
# (1,2) Maximum deficit volume (VDmax) and maximum deficit duration (DDmax) for Q < Qlow
VD <- 0.0; VDmax <- 0.0
DD <-0; DDmax <- 0
Qlow <- 0.5*mean(Qmon)   # threshold
for (i in 1:nmo_obs) {
  if (Qmon[i] < Qlow)   {VD <- VD + (Qlow-Qmon[i]); DD <- DD+1}
  else {VDmax <- max(VD,VDmax); VD <- 0; DDmax <- max(DD,DDmax); DD <- 0}
}
VDmax <- VDmax*2.628   # scale factor to obtain hm^3 from sum of m^3/s over several months (2.628*10^6 s/mon)

# (3,4) Maximum surplus volume (VSmax) and maximum surplus duration (DSmax) for Q > Qhigh
VS <- 0.0; VSmax <- 0.0
DS <-0; DSmax <- 0
Qhigh <- 2.0*mean(Qmon)   # threshold
for (i in 1:nmo_obs) {
  if (Qmon[i] > Qhigh)   {VS <- VS + (Qmon[i]-Qhigh); DS <- DS+1}
  else {VSmax <- max(VS,VSmax); VS <- 0; DSmax <- max(DS,DSmax); DS <- 0}
}
VSmax <- VSmax*2.628   # scale factor to obtain hm^3 from sum of m^3/s over several months (2.628*10^6 s/mon)

# (5,6) Storage capacity for total compensation of flows (SM) and longest dry period (LDP) 
#Diff <- numeric(length=length(Qmon))
Diff <- cumsum(Qmon - mean(Qmon))
SM <- max(Diff)-min(Diff)
SM <- SM*2.628                                 # scale factor to obtain hm^3 
LDP <- abs(which.max(Diff)-which.min(Diff))    # get index for max and min

# Calculate storage statistics for SIM flow------------------------------------>
# (1,2) Maximum deficit volume (VDmax1) and maximum deficit duration (DDmax1) for Q < Qlow
Qlow <- 0.5*mean(Qmon)   # absolute threshold fixed from OBS !!!
for (j in 1:nrea) {
  VD <- 0.0; VDmax1[j] <- 0.0
  DD <-0; DDmax1[j] <- 0
  for (i in 1:nmo_sim)
  {
    if (Qmon1[i,j] < Qlow)   {VD <- VD + (Qlow-Qmon1[i,j]); DD <- DD+1}
    else {VDmax1[j] <- max(VD,VDmax1[j]); VD <- 0; DDmax1[j] <- max(DD,DDmax1[j]); DD <- 0}
  }
  VDmax1[j] <- VDmax1[j]*2.628   # scale factor to obtain hm^3 from sum of m^3/s over several months (2.628*10^6 s/mon) 
}

# (3,4) Maximum surplus volume (VSmax) and maximum surplus duration (DSmax) for Q > Qhigh
Qhigh <- 2.0*mean(Qmon)   # absolute threshold fixed from OBS !!!
for (j in 1:nrea) {
  VS <- 0.0; VSmax1[j] <- 0.0
  DS <-0; DSmax1[j] <- 0
  for (i in 1:nmo_sim)
  {
    if (Qmon1[i,j] > Qhigh)   {VS <- VS + (Qmon1[i,j]-Qhigh); DS <- DS+1}
    else {VSmax1[j] <- max(VS,VSmax1[j]); VS <- 0; DSmax1[j] <- max(DS,DSmax1[j]); DS <- 0}
  }
  VSmax1[j] <- VSmax1[j]*2.628   # scale factor to obtain hm^3 from sum of m^3/s over several months (2.628*10^6 s/mon) 
}

# (5,6) Storage capacity for total compensation of flows (SM) and longest dry period (LDP) 
for (j in 1:nrea) {
  Diff1[,j] <- cumsum(Qmon1[,j] - mean(Qmon1[,j]))
  SM1[j] <- max(Diff1[,j])-min(Diff1[,j])
  SM1[j] <- SM1[j]*2.628                                      # scale factor to obtain hm^3 
  LDP1[j] <- abs(which.max(Diff1[,j])-which.min(Diff1[,j]))   # get index for max and min
}

# Plot storage capacities
# Box-Plot StDev monthly Q
sc[,1] <- VDmax1/VDmax
sc[,2] <- DDmax1/DDmax
sc[,3] <- VSmax1/VSmax
sc[,4] <- DSmax1/DSmax
sc[,5] <- SM1/SM
sc[,6] <- LDP1/LDP

boxplot(sc,  range=0, col = "lightblue", names=c('VD','DD','VS','DS','SM','LDP'), xlab = ' ', ylab='SIM/OBS [-]')   # grouped by month
points(rep(1.0,6),pch=4,col="red",lwd=2)         # add obs mean value as red cross

plot(Qacf_su, type='n', xlab='lag', ylab='ACF Summer [-]', ylim=c(-1.0,1.0), xlim=c(1,15))
#apply(Qacf1_su,1,lines,col='grey')
for(i in 1:nrow(Qacf1_su)) lines(Qacf1_su[i,],col='grey')
lines(Qacf_su,col='red',lwd=2)
lines(apply(Qacf1_su,2,mean),col='black',lwd=2,lty=2)
abline(a=0, b=0, col="black",lty=2)

plot(Qacf_wi, type='n', xlab='lag', ylab='ACF Winter [-]', ylim=c(-1.0,1.0), xlim=c(1,15))
#apply(Qacf1_wi,1,lines,col='grey')
for(i in 1:nrow(Qacf1_wi)) lines(Qacf1_wi[i,],col='grey')
lines(Qacf_wi,col='red',lwd=2)
lines(apply(Qacf1_wi,2,mean),col='black',lwd=2,lty=2)
abline(a=0, b=0, col="black",lty=2)

#if(flag.save) dev.off()

par(op)
```

## IQR - monthly 

```{r IQR-mon}
# Calculate percentage of cases where OBS is covered by IQR of SIM 
# over 6 statistics and 12 months (6*12)
Qmet_obs <- list(Qm, Qs, Qmin, Qmax, Qsk, Qcor)
Qmet_sim <- list(Qm1, Qs1, Qmin1, Qmax1, Qsk1, Qcor1)
n_met <- length(Qmet_obs)
theta1 <- 0.25; theta2 <- 0.75
theta3 <- 0.01; theta4 <- 0.99

IQR_cov <- 0
for(i in 1:12) {
  for(j in 1:n_met) {
    Q_tmp <- Qmet_obs[[j]]
    Q_tmp1<- Qmet_sim[[j]]
  
    if(Q_tmp[i] >= quantile(Q_tmp1[,i],theta1) & Q_tmp[i] <= quantile(Q_tmp1[,i],theta2)) IQR_cov <- IQR_cov+1

    if((Q_tmp[i] <= quantile(Q_tmp1[,i],theta1) & Q_tmp[i] >= quantile(Q_tmp1[,i],theta3)) ||
       (Q_tmp[i] <= quantile(Q_tmp1[,i],theta4) & Q_tmp[i] >= quantile(Q_tmp1[,i],theta2))) IQR_cov <- IQR_cov+0.5
    
    #if(Q_tmp[i] >= quantile(Q_tmp1[,i],theta1) & Q_tmp[i] <= quantile(Q_tmp1[,i],theta2)) IQR_cov <- IQR_cov+1
  }
}

IQR_cov <- IQR_cov/(n_met*12)

print(paste0('IQR_cov=',round(IQR_cov,3)))
```

## Monthly Climatology 

```{r fig-mon-clim, fig.width=9, fig.height=7}
## monthly climatology
summary(Qsim_mon[[1]]) 
summary(Qmon_obs)

#monthly average across years
x.mon.mean_r <- Qmon_sim[,.(sim=mean(sim), obs=mean(obs)),by=c("month","group")]
#summary(x.mon.mean_r)

x.mon.obs <- aggregate(Qm~month, Qmon_obs, FUN=mean) 
#summary(x.mon.obs)

p.mon.clim1 <- ggplot(x.mon.mean_r,aes(x=factor(month), y=sim)) +

  geom_boxplot() +
  stat_summary(fun=mean, geom="point", shape=20, size=3, color="black") +

  geom_point(data=x.mon.obs, aes(x=factor(month), y=Qm, color="Qobs_all"),shape=19) +
  geom_point(data=x.mon.mean_r, aes(x=factor(month), y=obs, color="Qobs_val"),shape=17) +

  scale_y_continuous(breaks = pretty_breaks()) +
  scale_color_manual(values=c("red","blue")) +

  labs(x="Month",y="Streamflow",title="Monthly Climatology",
       color=NULL, fill=NULL) +
  guides(fill="none") +
  theme_bw() +
  theme(text = element_text(size = 16,face='bold',family="serif"),
        plot.margin = unit(c(0.1,0.1,0.1, 0.1), "cm"),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),

        # strip.background = element_blank(),
        # strip.placement = "outside",

        legend.position = "right",
        #legend.position = c(.1,.85),
        legend.direction = c("horizontal","vertical")[2],
        legend.background = element_rect(fill="transparent"),
        legend.title=element_blank()
  )

p.mon.clim1 %>% print()
```

## Seasonal Climatology 

```{r fig-sea-clim, fig.width=9, fig.height=7}
#average across years
x.ses.mean_r <- Qmon_sim[,.(sim=mean(sim), obs=mean(obs)),by=c("season","group")]
#summary(x.mon.mean_r)

p.sea.clim1 <- ggplot(x.ses.mean_r,aes(x=factor(season), y=sim)) +

  geom_boxplot() +
  stat_summary(fun=mean, geom="point", shape=20, size=3, color="black") +

  geom_point(data=x.ses.mean_r, aes(x=factor(season), y=obs, color="Qobs"),shape=17) +

  scale_y_continuous(breaks = pretty_breaks()) +
  scale_color_manual(values="red") +

  labs(x="Season",y="Streamflow",title="Seasonal Climatology",
       color=NULL, fill=NULL) +
  guides(fill="none") +
  theme_bw() +
  theme(text = element_text(size = 16,face='bold',family="serif"),
        plot.margin = unit(c(0.1,0.1,0.1, 0.1), "cm"),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),

        # strip.background = element_blank(),
        # strip.placement = "outside",

        legend.position = "right",
        #legend.position = c(.1,.85),
        legend.direction = c("horizontal","vertical")[2],
        legend.background = element_rect(fill="transparent"),
        legend.title=element_blank()
  )

p.sea.clim1 %>% print()
```

# Yearly

## Caculate statistics

```{r ann}
#------------------------------------------------------------------------#
# OBS input
Qyr <- Qyr_obs$Qa[ind_yrs]
Qyr1 <- Qmon_sim[,.(value=mean(sim)),by=c("year","group")] %>% spread(group, value) %>% 
  dplyr::select(!c(year)) %>% as.matrix()
#summary(Qyr1)
#------------------------------------------------------------------------#
# OBS initialisations
Qm <- numeric(length=1)    # mean
Qs <- numeric(length=1)    # stddev
Qmin <- numeric(length=1)
Qmax <- numeric(length=1)
Qsk <- numeric(length=1)  # skewness
Qacf <- numeric(length=15)  # ACF 15 lags

# SIM initialisations
Qm1 <- numeric(length=nrea)    # mean for each realisation
Qs1 <- numeric(length=nrea)    # stddev
Qmin1 <- numeric(length=nrea)  # min
Qmax1 <- numeric(length=nrea)  # max
Qsk1 <- numeric(length=nrea)   # skewness
Qacf1 <- array(NA, dim=c(nrea,15))   # ACF 15 lags
amin <- numeric(length=15)  # min ACF 15 lags
amax <- numeric(length=15)  # max ACF 15 lags
bpv <- array(NA, dim=c(nrea,4)) # array to hold box-plot-vars cs (dim: nreas, 4 vars)

#------------------------------------------------------------------------#
# Basic OBS statistics
Qm <- mean(Qyr)
Qs <- sd(Qyr)
Qmin <- min(Qyr)
Qmax <- max(Qyr)
Qsk <- skewness(Qyr)  # req. package moments; method not clear !

# Basic SIM statistics for each realisation
for (j in 1:nrea) {
  Qm1[j] <- mean(Qyr1[,j])
  Qs1[j] <- sd(Qyr1[,j])
  Qmin1[j] <- min(Qyr1[,j])
  Qmax1[j] <- max(Qyr1[,j])
  Qsk1[j] <- skewness(Qyr1[,j])  # req. package moments; method not clear !  
}

# ACF OBS & SIM lags
Qacf <- acf(Qyr, plot=FALSE)$acf[2:16]
for(j in 1:nrea) { Qacf1[j,1:15] <- acf(Qyr1[,j], plot=FALSE)$acf[2:16]}
```

## Basic statistics - yearly

```{r fig-yr-stat, fig.width=9, fig.height=7}
#if(flag.save) jpeg("Qyr_valid.jpg", units="cm", width=17, height=13, res=dpi_fig)

# plot layout as matrix filled by cols
layout(matrix(c(1:4), nrow=2, ncol=2, byrow=TRUE))    
par(ps="12",              # size
    las=1,                # axis label orientation
    mar=c(2,5,1,2),       # no. of plot margin lines
    oma=c(2,2,2,1),       # no. of outer margin lines
    mgp=c(2.5,0.5,0),         # margin line for the axis title, labels and line.
    pty="m")              # maximum plot region used


# Plot OBs annual TS
plot(Qyr, type='s',lwd=2,col='blue', xlab='Zeit [Jahre]', ylab='Qyr OBS [m3/s]')    # plot annual time seris
abline(a=mean(Qyr), b=0, col="black",lty=2)

# Plot 1 realisation of SIM annual TS
plot(Qyr1[,1], type='s',lwd=2,col='blue', xlab='Zeit [Jahre]', ylab='Qyr SIM [m3/s]')    # plot annual time seris
#line(Qyr1[,2], type='l', lwd=1, col='red')
abline(a=mean(Qyr1[,2]), b=0, col="black",lty=2)

# Combine variables for Box-Plot; build ratios SIM/OBS 
bpv[,1] <- Qm1/Qm
bpv[,2] <- Qs1/Qs
bpv[,3] <- Qmin1/Qmin
bpv[,4] <- Qmax1/Qmax
#bpv[,5] <- Qsk1/Qsk

# Box-plot annual statistics
boxplot(bpv,  range=0, col = "lightblue", names=c('mean','SD','min','max'), 
        xlab = ' ', ylab='SIM/OBS [-]')   # grouped by month
points(rep(1.0,5),pch=4,col="red",lwd=2)         # add obs mean value as red cross

# Plot ACF  -- lines
plot(Qacf, type='n', xlab='lag', ylab='ACF [-]', ylim=c(-1.0,1.0), xlim=c(1,14))
for(i in 1:nrow(Qacf1)) lines(Qacf1[i,],col='grey')
lines(Qacf,col='red',lwd=2)
lines(apply(Qacf1,2,mean),col='black',lwd=2,lty=2)

#if(flag.save) dev.off()
```

# Daily 

## Caculate statistics

```{r day}
#------------------------------------------------------------------------#
# OBS input
Qday <- Qday_obs$Qd[ind_day]
mon <- Qday_obs$mon[ind_day]
summary(Qday)
# SIM input
Qday1 <- sapply(Qsim_day, function(ls) ls$Qd)
mon1 <- mon
#summary(Qday1)

#------------------------------------------------------------------------#
# OBS initialisations
Qm <- numeric(length=12)                      # mean
Qs <- numeric(length=12)                      # stddev
Qmin <- numeric(length=12)
Qmax <- numeric(length=12)
Qsk <- numeric(length=12)                     # skewness
Qcor <- numeric(length=12)                    # lag-1 autocorrelation
Qacf_su <- numeric(length=30)                 # ACF summer 30 lags
Qacf_wi <- numeric(length=30)                 # ACF winter 30 lags

# SIM initialisations
Qm1 <- array(NA, dim=c(nrea,12))    # mean (dim: nrea realisations, 12 months)
Qs1 <- array(NA, dim=c(nrea,12))    # stddev
Qmin1 <- array(NA, dim=c(nrea,12))  # min
Qmax1 <- array(NA, dim=c(nrea,12))  # max
Qsk1 <- array(NA, dim=c(nrea,12))   # skewness
Qcor1 <- array(NA, dim=c(nrea,12))   # lag-1 autocorrelation
Qacf1_su <- array(NA, dim=c(nrea,30))   # ACF summer 30 lags
Qacf1_wi <- array(NA, dim=c(nrea,30))   # ACF summer 30 lags
amin <- numeric(length=30)  # min ACF 30 lags
amax <- numeric(length=30)  # max ACF 30 lags
fmin <- numeric(length=nyr_sim*31)  # min FDC by month
fmax <- numeric(length=nyr_sim*31)  # max FDC by month
Qsrt1 <- array(NA, dim=c(nyr_sim*31,nrea))  # sorted Qday1 by month

#------------------------------------------------------------------------#
# Basic OBS statistics by month
for (i in 1:12)  {
  Qm[i] <- mean(Qday[mon == i])
  Qs[i] <- sd(Qday[mon == i])
  Qmin[i] <- min(Qday[mon == i])
  Qmax[i] <- max(Qday[mon == i])
  Qsk[i] <- skewness(Qday[mon == i])  # req. package moments; method not clear !
}

# Basic SIM statistics by month for each realization
for (i in 1:12) {
  for (j in 1:nrea){
    Qm1[j,i] <- mean(Qday1[,j][mon1 == i])
    Qs1[j,i] <- sd(Qday1[,j][mon1 == i])
    Qmin1[j,i] <- min(Qday1[,j][mon1 == i])
    Qmax1[j,i] <- max(Qday1[,j][mon1 == i])
    Qsk1[j,i] <- skewness(Qday1[,j][mon1 == i])  # req. package moments; method not clear !  
  }
}

# Lag-1 OBS autocorrelation per month
for (i in 1:12) {
  Qcor[i] <- acf(Qday[mon == i], plot=FALSE, lag.max=1)$acf[2]
}

# Lag-1 SIM autocorrelation per month
for (i in 1:12) {
  for (j in 1:nrea) {
    Qcor1[j,i] <- acf(Qday1[,j][mon1 == i], plot=FALSE, lag.max=1)$acf[2]    
  }
}


# ACF OBS & SIM 30 lags for summer and winter
Qacf_su <- acf(Qday[mon>4 & mon<11], plot=FALSE)$acf[1:30]    # OBS summer
Qacf_wi <- acf(Qday[mon<5 | mon>10], plot=FALSE)$acf[1:30]    # OBS winter

for (j in 1:nrea) { 
  Qacf1_su[j,1:30] <- acf(Qday1[,j][mon1>4 & mon1<11], plot=FALSE)$acf[1:30] 
}# SIM nrea reas summer

for (j in 1:nrea) { 
  Qacf1_wi[j,1:30] <- acf(Qday1[,j][mon1<5 | mon1>10], plot=FALSE)$acf[1:30] 
}    # SIM nrea reas winter

```

## Basic statistics - daily

```{r fig-day-stat, warning=FALSE}
#if(flag.save) jpeg("Qday1valid.jpg", units="cm", width=17, height=17, res=dpi_fig)

# plot layout as matrix filled by cols
layout(matrix(c(1:6), nrow=3, ncol=2, byrow=TRUE))    
par(ps="12",              # size
    las=1,                # axis label orientation
    mar=c(2,5,1,2),       # no. of plot margin lines
    oma=c(2,2,2,1),       # no. of outer margin lines
    mgp=c(3,1,0),         # margin line for the axis title, labels and line.
    pty="m")              # maximum plot region used

# Box-Plot mean daily Q
boxplot(Qm1, range=0, col = "lightblue", xlab='Month', ylab='Average [m3/s]')   # grouped by month
points(Qm,pch=4,col="red",lwd=2)         # add obs mean value as red cross

# Box-Plot Std Dev daily Q
boxplot(Qs1, range=0, col = "lightblue", #ylim=c(0,100),
        xlab='Month', ylab='Std. Dev. [m3/s]')   # grouped by month
points(Qs,pch=4,col="red",lwd=2)         # add obs mean value as red cross


# Box-Plot Min daily Q
boxplot(Qmin1, range=0, col = "lightblue", #ylim=c(0,20),
        xlab='Month', ylab='Min [m3/s]')   # grouped by month
points(Qmin,pch=4,col="red",lwd=2)         # add obs mean value as red cross


# Box-Plot Max daily Q
boxplot(Qmax1, range=0, col = "lightblue", #ylim=c(0,2000),
        xlab='Month', ylab='Max [m3/s]')   # grouped by month
points(Qmax,pch=4,col="red",lwd=2)         # add obs mean value as red cross


# Box-Plot Schiefe daily Q
boxplot(Qsk1, range=0, col = "lightblue", #ylim=c(0,20),
        xlab='Month', ylab='Schiefe [-]')   # grouped by month
points(Qsk,pch=4,col="red",lwd=2)         # add obs mean value as red cross


# Box-Plot Lag-1 autocorrelation daily Q
boxplot(Qcor1, range=0, col = "lightblue", #ylim=c(0.5,1),
        xlab='Month', ylab='Lag-1 Corr [-]')   # grouped by month
points(Qcor,pch=4,col="red",lwd=2)         # add obs mean value as red cross


#if(flag.save) dev.off()
```

## IQR - daily

```{r IQR-day}
# Calculate percentage of cases where OBS is covered by IQR of SIM 
# over 6 statistics and 12 months (6*12)
Qmet_obs <- list(Qm, Qs, Qmin, Qmax, Qsk, Qcor)
Qmet_sim <- list(Qm1, Qs1, Qmin1, Qmax1, Qsk1, Qcor1)
n_met <- length(Qmet_obs)
theta1 <- 0.25; theta2 <- 0.75
theta3 <- 0.01; theta4 <- 0.99

IQR_cov <- 0
for(i in 1:12) {
  for(j in 1:n_met) {
    Q_tmp <- Qmet_obs[[j]]
    Q_tmp1<- Qmet_sim[[j]]
  
    if(Q_tmp[i] >= quantile(Q_tmp1[,i],theta1) & Q_tmp[i] <= quantile(Q_tmp1[,i],theta2)) IQR_cov <- IQR_cov+1

    if((Q_tmp[i] <= quantile(Q_tmp1[,i],theta1) & Q_tmp[i] >= quantile(Q_tmp1[,i],theta3)) ||
       (Q_tmp[i] <= quantile(Q_tmp1[,i],theta4) & Q_tmp[i] >= quantile(Q_tmp1[,i],theta2))) IQR_cov <- IQR_cov+0.5
    
    #if(Q_tmp[i] >= quantile(Q_tmp1[,i],theta1) & Q_tmp[i] <= quantile(Q_tmp1[,i],theta2)) IQR_cov <- IQR_cov+1
  }
}

IQR_cov <- IQR_cov/(n_met*12)

print(paste0('IQR_cov=',round(IQR_cov,3)))
```

## CDF

```{r fig-day-cdf, warning=FALSE}
# plot summer & winter ACF and FDC for 4 selected months as 3 x 2 matrix 
#if(flag.save) jpeg("Qday2valid.jpg", units="cm", width=17, height=17, res=400)

# plot layout as matrix filled by cols
layout(matrix(c(1:6), nrow=3, ncol=2, byrow=TRUE))    
par(ps="12",              # size
    las=1,                # axis label orientation
    mar=c(2,5,1,2),       # no. of plot margin lines
    oma=c(2,2,2,1),       # no. of outer margin lines
    mgp=c(3,1,0),         # margin line for the axis title, labels and line.
    pty="m")              # maximum plot region used

# Plot daily ACF summer - polygon
x <- c(1:30)
xx <- c(x,rev(x))   # x-values for polygon
for (k in 1:30) {
  amin[k] <- min(Qacf1_su[,k])
  amax[k] <- max(Qacf1_su[,k])
}
yy <- c(amin,rev(amax))   # y-values for polygon
plot(Qacf_su, type='l', ylim=c(0,1), xlab='lag', ylab='ACF Summer [-]')
polygon(xx,yy,col="lightgrey",border=NA)
lines(Qacf_su, type='l')

# Plot daily ACF winter - polygon
x <- c(1:30)
xx <- c(x,rev(x))   # x-values for polygon
for (k in 1:30) {
  amin[k] <- min(Qacf1_wi[,k])
  amax[k] <- max(Qacf1_wi[,k])
}
yy <- c(amin,rev(amax))   # y-values for polygon
plot(Qacf_wi, type='l', ylim=c(0,1), xlab='lag', ylab='ACF Winter [-]')
polygon(xx,yy,col="lightgrey",border=NA)
if(nensemble==1) lines(xx,yy,col="red",border=NA)
lines(Qacf_wi, type='l')

# Calculate and plot flow duration curve by month OBS & SIM
ystr<-c('Jan','Feb','Mar','Apr','May','Jun','Jul','Aug','Sep','Oct','Nov','Dec')

# months January, April, July, October
for (m in seq(1,12,by=3)) {
  Qsrt <- sort(Qday[mon==m])
  n0 <- length(Qday[mon==m])
  Fn <- seq(1,n0)/(n0+1)
  n1 <- length(Qday1[,1][mon1==m])
  for (j in 1:nrea) { Qsrt1[1:n1,j] <- sort(Qday1[,j][mon1==m]) }
  Fn1 <- seq(1, n1)/(n1+1)
  xx <- c(Fn1,rev(Fn1))
  for (k in 1:n1)
  {
    fmin[k] <- min(Qsrt1[k,])
    fmax[k] <- max(Qsrt1[k,])
  }
  yy <- c(fmin[1:n1],rev(fmax[1:n1]))   # y-values for polygon
  plot(Fn, Qsrt, type='l', pch=1, cex=0.5, log="y", #ylim=c(1,500),
       xlab='Pu [-]', ylab=paste('Q [m3/s] - ',ystr[m]))
  polygon(xx,yy,col="lightgrey",border=NA)
  lines(Fn,Qsrt, type='l', pch=1, cex=0.5)
}

#if(flag.save) dev.off()
```

# Seasonality change 
```{r seasonality}
out.Qmon <- list()
gcms <- c("hist","rcp85_near", "rcp85_far")
for(i in 1:length(gcms)) {
  flag.gcm <- gcms[i]
  print(flag.gcm)
  Qsim_mon <- get(load(paste0(path.out, "Harz_Qmon_",flag.sel,"_r",nensemble,"_",flag.gcm,flag.ver,".Rdat"))) 
  
  out.Qmon[[i]] <- rbindlist(Qsim_mon, idcol="group") %>% data.frame()
}
names(out.Qmon) <- gcms

summary(out.Qmon[[3]])

x.mon.mean_r <- rbindlist(out.Qmon, idcol="gcm") %>% rename("sim"="Qm")
x.mon.mean_r$gcm <- factor(x.mon.mean_r$gcm)

x.mon.mean_r$gcm <- factor(x.mon.mean_r$gcm, levels=gcms)
summary(x.mon.mean_r)

#monthly average across years
x.mon.mean_r1 <- x.mon.mean_r[,.(sim=mean(sim)),by=c("month","group","gcm")]

summary(x.mon.mean_r1)

p.mon.clim1 <- ggplot(x.mon.mean_r1, aes(x=factor(month), y=sim, fill=gcm)) +

  geom_boxplot() +
  #stat_boxplot(geom ='errorbar')+
  #stat_summary(fun=mean, geom="point", shape=20, size=3, color="black") +
  
  scale_y_continuous(breaks = pretty_breaks()) +
  scale_fill_manual(values=c("grey","blue","red")) +

  labs(x="Month",y="Qmon[m3/s]",#title="Monthly Climatology",
       color=NULL, fill=NULL) +
  #guides(fill="none") +

  theme_bw() +
  theme(text = element_text(size = 16,face='bold',family="serif"),
        plot.margin = unit(c(0.1,0.1,0.1, 0.1), "cm"),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),

        # strip.background = element_blank(),
        # strip.placement = "outside",

        #legend.position = "right",
        legend.position = c(.6,.9),
        legend.direction = c("horizontal","vertical")[1],
        legend.background = element_rect(fill="transparent"),
        legend.title=element_blank()
  )

p.mon.clim1 %>% print()
```

# Storage change 

```{r storage, fig.height = 9, fig.width = 12}
# Calculate storage statistics for current flows----------------------------------->
# (1,2) Maximum deficit volume (VDmax) and maximum deficit duration (DDmax) for Q < Qlow
Qmon <- Qmon_obs$Qm[ind_mon]
Qsim_mon <- out.Qmon[[1]]
summary(Qmon)
Qmon1 <- dplyr::select(Qsim_mon,!nn) %>% spread(group, Qm) %>% dplyr::select(!c(year,month)) %>%
  as.matrix()
s.hist <- storage(Qmon, Qmon1)

out.storage <- list()
for(i_gcm in 2:length(gcms)) {
  flag.gcm <- gcms[i_gcm]
  print(flag.gcm)
  
  Qsim_mon <- out.Qmon[[i_gcm]]
  #summary(Qmon)
  Qmon1 <- dplyr::select(Qsim_mon,!nn) %>% spread(group, Qm) %>% dplyr::select(!c(year,month)) %>%
    as.matrix()

  sc <- array(NA, dim=c(nrea,6))    # storage characteristics (dim: 'nrea' realisations, 6 characteristics)


  # Calculate storage statistics for SIM flow------------------------------------>
  s.gcm <- storage(Qmon, Qmon1)
  VDmax1 <- s.gcm$VD
  DDmax1 <- s.gcm$DD
  VSmax1 <- s.gcm$VS
  DSmax1 <- s.gcm$DS  
  SM1<- s.gcm$SM  
  LDP1 <- s.gcm$LDP

  # Plot storage capacities
  # Box-Plot StDev monthly Q
  sc[,1] <- VDmax1/s.hist$VD
  sc[,2] <- DDmax1/s.hist$DD
  sc[,3] <- VSmax1/s.hist$VS
  sc[,4] <- DSmax1/s.hist$DS
  sc[,5] <- SM1/s.hist$SM
  sc[,6] <- LDP1/s.hist$LDP

  sc.df <- data.frame(sc)
  colnames(sc.df) <- c("VD","DD","VS","DS","SM","LDP")
  out.storage[[i_gcm]] <- sc.df

}
names(out.storage) <- gcms

sc.plot <- rbindlist(out.storage,idcol="gcm")
sc.plot$gcm <- factor(sc.plot$gcm)
sc.plot$gcm <- factor(sc.plot$gcm, levels=gcms)
#summary(sc.plot)

sc.plot1 <- sc.plot %>% gather(group,sc,2:7)
sc.plot1$group <- factor(sc.plot1$group)
sc.plot1$group <- factor(sc.plot1$group, levels= c("VD","DD","VS","DS","SM","LDP"))

summary(sc.plot1)
p.sc <- ggplot(subset(sc.plot1,gcm!="hist"), aes(x=group, y=sc, color=gcm)) +

  geom_boxplot(position = position_dodge(width = 1)) +
  #stat_boxplot(geom ='errorbar')+
  stat_summary(fun=mean, geom="point", shape=17, size=5,position = position_dodge(width = 1)) +

  geom_abline(slope=0,intercept=1, color="black") +

  scale_y_continuous(limits=c(0,2), breaks = pretty_breaks()) +
  scale_color_manual(values=c("grey","blue","red")[2:3]) +

  labs(x="Storage characteristics",y="Future/Past[-]",
       color=NULL, fill=NULL) +

  theme_bw() +
  theme(text = element_text(size = 16,face='bold',family="serif"),
        plot.margin = unit(c(0.1,0.1,0.1, 0.1), "cm"),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),

        # strip.background = element_blank(),
        # strip.placement = "outside",

        #legend.position = "right",
        legend.position = c(.2,.9),
        legend.direction = c("horizontal","vertical")[1],
        legend.background = element_rect(fill="transparent"),
        legend.title=element_blank()
  )
p.sc

fig <- cowplot::plot_grid(p.mon.clim1,p.sc, ncol=2, #labels=p.labels,
                          label_size=12,
                          label_fontfamily = "serif",
                          label_fontface = "bold",
                          label_colour = "black")
#fig 
```

